Dependency in Refinement Logics

### 2015-10-17 doc/paper.tex
Add constructible subgoals section

\documentclass[11pt]{amsart}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath, stmaryrd, scalerel}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{unicode-math}
\usepackage{xltxtra}
\usepackage{xunicode}

@@

% fonts
\setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
\setmathfont{Asana Math}

@@

\newcommand\Member[2]{\InputMode{#1}\in\InputMode{#2}}
\newcommand\IdTac{\mathtt{id}}
\newcommand\FailTac{\mathtt{fail}}
\newcommand\OrElseTac[2]{#1\vert#2}
\newcommand\ThenTac[2]{#1;#2}
\newcommand\ThenLTac[2]{#1;[#2]}
\newcommand\ProdIntroRule{\mathtt{ProdIntro}}
\newcommand\DProdIntroRule[1]{\mathtt{DepProdIntro}\{#1\}}
\newcommand\DProdIntroRuleNullary{\mathtt{DepProdIntro}}
\newcommand\TyProd[2]{#1 \times #2}
\newcommand\TyDProd[3]{(#2:#1) \times #3}
\newcommand\TyMember[2]{#1\in#2}
\newcommand\Dom[1]{\left\vert#1\right\vert}
\newcommand\SubsetEq[2]{\InputMode{#1}\subseteq\InputMode{#2}}
\newcommand\FV[1]{\textit{FV}\left(#1\right)}
\newcommand\Subst[3]{\left[#1\;/\;#2\right]#3}
\newcommand\ModLCF{\textbf{\emph{Modernized LCF}}}
\newcommand\DepLCF{\textbf{\emph{Dependent LCF}}}

@@

\title{Dependency in Refinement Logics}

\author{Jonathan Sterling}
\address{}
\maketitle

@@

\onehalfspacing

@@

\section{modernized lcf: the logic of tactics}

@@

>>>>

\documentclass[11pt]{article}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath, amsthm, stmaryrd, scalerel}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{unicode-math}
\usepackage{xltxtra}
\usepackage{xunicode}

@@

% fonts
% \setromanfont[Mapping=tex-text, Ligatures=Rare, Numbers=OldStyle]{Linux Libertine O}
\setmathfont{Asana Math}

@@

\newcommand\Member[2]{\InputMode{#1}\in\InputMode{#2}}
\newcommand\MemberUnmoded[2]{#1\in #2}
\newcommand\IsTrue[1]{\InputMode{#1}\ \textit{true}}
\newcommand\IsTrueUnmoded[1]{#1\ \textit{true}}
\newcommand\IdTac{\mathtt{id}}
\newcommand\FailTac{\mathtt{fail}}
\newcommand\OrElseTac[2]{#1\vert#2}
\newcommand\ThenTac[2]{#1;#2}
\newcommand\ThenLTac[2]{#1;[#2]}
\newcommand\ProdIntroRule{\mathtt{ProdIntro}}
\newcommand\DProdIntroRule[1]{\mathtt{DepProdIntro}\{#1\}}
\newcommand\DProdIntroRuleNullary{\mathtt{DepProdIntro}}
\newcommand\TyProd[2]{#1 \times #2}
\newcommand\TyDProd[3]{(#2:#1) \times #3}
\newcommand\TyMember[2]{#1\in#2}
\newcommand\Dom[1]{\left\vert#1\right\vert}
\newcommand\SubsetEq[2]{\InputMode{#1}\subseteq\InputMode{#2}}
\newcommand\FV[1]{\textit{FV}\left(#1\right)}
\newcommand\Subst[3]{\left[#1\;/\;#2\right]#3}
\newcommand\ModLCF{\textbf{\emph{Modernized LCF}}}
\newcommand\DepLCF{\textbf{\emph{Dependent LCF}}}

@@

\title{Dependency in Refinement Logics}
\date{}
\author{Jonathan Sterling}
\maketitle

@@

@@

\section{\ModLCF: the logic of tactics}

@@

\section{\ModLCF\ and the constructible subgoals property}

An LCF-style refiner has a property called \emph{constructible subgoals}, which
means that the subgoals incurred by a rule or tactic may be constructed
independently, using only the statement of the main goal. One unfortunate
consequence of this principle is that it is impossible to define a refinement
rule which expresses a dependency between subgoals. The canonical example is
the introduction rule for dependent pairs in Type Theory:
\[
  \infer{
    \Member{\Tuple{M,N}}{\TyDProd{A}{x}{B}}
  }{
    \Member{M}{A} &
    \Member{N}{\Subst{M}{x}{B}}
  }
\]

However, consider what a refinement rule for this would look like; we would
like to take the goal $\IsTrue{\TyDProd{A}{x}{B}}$ to two subgoals (one for
each conjunct); the first subgoal $\IsTrue{A}$ is clear enough, but it is not
possible to even write down the second subgoal until we know the synthesis of
the first one.

Because a tactic in \ModLCF\ produces only a \emph{context} of independent
subgoals, we cannot give a refinement treatment to this subgoal, and must
instead write a family of refinement rules $\DProdIntroRule{w}$ fibred over
witnesses of the left conjunct $w$:

\[
  \Refine{\IsTrueUnmoded{\TyDProd{A}{x}{B}}}{\DProdIntroRule{w}}{%
    \left\{\begin{array}{l}
      \SG{\alpha}{\MemberUnmoded{w}{A}}\\
      \SG{\beta}{\IsTrueUnmoded{\Subst{w}{x}{B}}}
    \end{array}\right.
  }{\Tuple{w,\beta}}
\]

This is clearly unsatisfactory, since it breaks the natural flow of proof
development, whereby multiple goals may be refined simultaneously without
committing in advance to a particular solution. However, a more palatable
refinement rule that allows $\IsTrue{A}$ to be demonstrated by refinement is
simply not expressible in \ModLCF, since the sense of the second subgoal cannot
be expressed except by referring to the synthesis of the first subgoal.
